### 2018-03-16 figures/conny_metric_key.tex
towards a connectivity metric cartoon

\documentclass[bigger]{beamer}
\usepackage[orientation=portrait, size=a0, scale=1.4]{beamerposter}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{tabularx}

\setbeamertemplate{navigation symbols}{}%remove navigation symbols
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\input{beamer-lake-fig/beamer-lake-fig.tex}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Connectivity Metric Key}
\begin{frame}
  \begin{center}

  % cd %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns} [c]     
  \begin{column}{0.3\textwidth}
    \begin{tikzpicture}[scale = 0.6]
        \draw [-,ultra thick, cyan] (2.2,2.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
        \draw [-,ultra thick, cyan] (2.2,2.8) to [out=90,in=0] (1.1,3.1) to [out=270,in=90] (0,3.4);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 3.1cm] plot (\x:0.4+rnd*0.2);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 3cm, xshift = 2cm] plot (\x:0.6+rnd*0.2);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
  \end{column}

  \begin{column}{0.3\textwidth}
    \begin{tikzpicture}[scale = 0.6]
        \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 5cm, xshift = 2cm] plot (\x:0.6+rnd*0.2);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
  \end{column}
  \end{columns}

  % ll %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[c]  
    \begin{column}{0.3\textwidth}
      \begin{tikzpicture}[scale = 0.6]
        % \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,ultra thick, cyan] (2,3) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (0.5,2.9) to [out=0,in=90] (1.1,1.9) to [out=270,in=90] (0,0);

        \draw [-,ultra thick, cyan] (-1,4) to [out=200,in=0] (-2,3) to [out=290,in=45] (-1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (-0.5,2.9) to [out=250,in=250] (-1.1,1.9) to [out=270,in=90] (0,0);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \begin{column}{0.3\textwidth}
      \begin{tikzpicture}[scale = 0.6]
        %   \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,ultra thick, cyan] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

    % la %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[c]  
    \begin{column}{0.3\textwidth}
      \begin{tikzpicture}[scale = 0.6]
        % \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,ultra thick, cyan] (2,3) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (0.5,2.9) to [out=0,in=90] (1.1,1.9) to [out=270,in=90] (0,0);

        \draw [-,ultra thick, cyan] (-1,4) to [out=200,in=0] (-2,3) to [out=290,in=45] (-1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (-0.5,2.9) to [out=250,in=250] (-1.1,1.9) to [out=270,in=90] (0,0);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \begin{column}{0.3\textwidth}
      \begin{tikzpicture}[scale = 0.6]
        %   \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,ultra thick, cyan] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

  \end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}