### 2018-05-11 scripts/analysis.Rmd
update fig and table captions from gdoc

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("01_prepdata.R")
knitr::read_chunk("03_viz.R")
knitr::read_chunk("04_tables.R")
knitr::read_chunk("99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r prep_nes_lagos, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r lake_characteristics_table, echo=FALSE, results='asis', message=FALSE, warning=FALSE}

```

```{r fig.cap="Diagram showing the inter-lake watershed (IWS) and network watershed (NWS) of three lakes. Here the IWS of lake 3 encompasses the IWS of lake 2 because of it is smaller than 10 ha small size but it does not encompass the IWS of lake 1 because it has an area of at least 10 ha. In contrast to the IWS boundaries, the NWS boundaries extend to the headwaters of the lake chain.", echo=FALSE}
knitr::include_graphics("../figures/iws_nws.pdf")
```

```{r fig.cap="Connectivity metric definitions and examples of high and low connectivity lakes.", echo=FALSE}
knitr::include_graphics("../figures/conny_metric_key_crop.pdf")
```

```{r cor_mat_hmap, echo=FALSE, warning=FALSE, fig.width=3.5, fig.height=3.2, fig.cap="Correlation between connectivity metrics and selected lake characteristics.", message=FALSE}

```

```{r global_vollenweider_viz, echo=FALSE, fig.cap="Residence time (yr) versus P retention for the NES dataset and the global model fit to the data as the median and central 95% interval estimates.", warning=FALSE, fig.width = 3.5, fig.height = 3.2}

```

```{r partition_vollenweider_viz, echo=FALSE, fig.cap="Residence time (yr) versus P retention (%) for the NES dataset and hierarchical model fits to the data as the median and central 95% interval estimates.  The green lines and symbols are the estimates from the lower of the two partition groups while the purple lines are estimates for the upper of the two partition groups (see Table 2).", fig.width = 4.8, fig.height=5, warning=FALSE}

```

```{r k_viz, echo=FALSE, message=FALSE, fig.cap="Distribution of the k parameter from the Vollenweider's equation in low and high connectivity partitions at the (A) IWS and (B) NWS scales. Green symbols indicate the lower of the two partition groups while purple symbols represent the higher of the two partition groups (see Table 2).", fig.width=5, fig.height=3.8, warning=FALSE}

```

```{r model_results_table, echo=FALSE, results='asis'}

```

```{r maps, echo=FALSE, fig.cap="Maps showing the locations of lake connectivity partitions. Green symbols indicate the lower of the two partition groups while purple symbols represent the higher of the two partition groups (see Table 2).", warning=FALSE, message=FALSE}

```

<!-- * Olin Lake has a completely undeveloped shoreline -->
<!-- * Madison Lake is impaired according to MPCA -->

```{r recursive_partitioning_viz, echo=FALSE, eval=FALSE}

```

```{r graphical_exploratory_analysis, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8.4, fig.cap="Exploratory plots"}

```

```{r graphical_exploratory_analysis_cont, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6.3, fig.cap="Exploratory plots cont"}

```

```{r graphical_exploratory_analysis_cont_cont, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6.3, fig.cap="Exploratory plots cont x 2"}

```

```{r conceptual_summary, echo=FALSE, fig.cap="Connectivity metric effects on lake processing.", eval=FALSE}
dt <- data.frame(
  rbind(
    c("closest lake distance", "lower", "higher", "lower", "higher", "lake-stream", "long"), 
    c("average link length", "lower", "higher", "lower", "higher", "stream", "long"), 
    c("lakeconnection", "DR_Stream", "higher", "lower", "higher", "lake", "long"), 
    c("upstream lake area", "lower", "higher", "lower", "higher", "lake", "lat"), 
    c("stream density", "higher", "higher", "lower", "higher", "stream", "lat"),
    c("baseflow", "higher", "higher", "lower", "higher", "stream", "lat"), 
    c("stream order ratio", "higher", "higher", "lower", "higher", "stream", "long")))

names(dt) <- c("metric", "direction", "k", "connectivity", "in_lake_processing", "axis", "connectivity type")

knitr::kable(dt)
```