### 2018-05-31 figures/07_cor_mat_hmap.Rmd

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../scripts/03_viz.R")
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
  source("../scripts/01.5_loaddata.R")
```

```{r 07_cor_mat_hmap, echo=FALSE, warning=FALSE, fig.width=3.5, fig.height=3.2, fig.cap="Correlation between connectivity metrics and selected lake characteristics.", message=FALSE}

```