### 2018-06-18 figures/conny_metric_key.tex
rm uplake area

\documentclass[bigger]{beamer}
\usepackage[orientation=portrait, size=a10, scale=1]{beamerposter}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{tabularx}

\setbeamertemplate{navigation symbols}{}%remove navigation symbols
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\input{beamer-lake-fig/beamer-lake-fig.tex}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Connectivity Metric Key}
\begin{frame}
  \begin{center}

    \begin{columns} [c]     
      \begin{column}{0.4\textwidth}        
      \end{column}

    \begin{column}{0.25\textwidth}
      \huge{Lower Connectivity}
    \end{column}

    \begin{column}{0.25\textwidth}
      \huge{Higher Connectivity}
    \end{column}
    \end{columns}

  % cd %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}    
    \begin{column}[b]{0.4\textwidth}
      \huge{\textbf{Closest Lake Distance:} Network distance to the closest upstream lake.}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
    \begin{tikzpicture}[scale = 2]
        \draw [-,ultra thick, cyan] (2.2,2.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 3cm, xshift = 2cm] plot (\x:0.6+rnd*0.2);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
    \begin{tikzpicture}[scale = 2]
        \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 5cm, xshift = 2cm] plot (\x:0.6+rnd*0.2);
        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
  \end{column}
  \end{columns}

  % ll %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns} 
    \begin{column}[b]{0.4\textwidth}
      \huge{\textbf{Average Link Length:} Sum of the total length of stream reaches between junctions divided by the total number of reaches.}
    \end{column}

    \begin{column}[b]{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        %\draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,ultra thick, cyan] (1.8,2.6) to [out=180,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (0.5,2.9) to [out=0,in=90] (1.1,1.9) to [out=270,in=90] (0,0);

        \draw [-,ultra thick, cyan] (-1.2,3.4) to [out=200,in=0] (-1.8,2.4) to [out=290,in=110] (-1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (-0.5,2.9) to [out=250,in=70] (-1.1,1.9) to [out=270,in=90] (0,0);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \begin{column}[b]{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
           %\draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,ultra thick, cyan] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw [-,ultra thick, cyan] (-2.2,4.8) to [out=90,in=180] (-1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,ultra thick, cyan] (-0.9,4) to [out=260,in=180] (-1.1,1.9) to [out=270,in=90] (0,0);

          \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

  % la %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  % \begin{columns}[b]  
  %   \begin{column}{0.4\textwidth}
  %     \huge{\textbf{Upstream Lake Area:} Total area of upstream lakes limited to lakes that are greater than or equal to 4ha.}
  %   \end{column}
  % 
  %   \begin{column}{0.25\textwidth}
  %     \begin{tikzpicture}[scale = 2]
  %         \draw [-,ultra thick, cyan] (2.2,3.7) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
  %         \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 3.5cm, xshift = 2cm] plot (\x:0.5+rnd*0.2);
  %         \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
  %       \end{tikzpicture}
  %   \end{column}
  % 
  %   \begin{column}{0.25\textwidth}
  %     \begin{tikzpicture}[scale = 2]
  %         \draw [-,ultra thick, cyan] (2.2,4.1) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
  %         \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan, yshift = 3.5cm, xshift = 2cm] plot (\x:0.9+rnd*0.2);
  %         \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
  %       \end{tikzpicture}
  %   \end{column}
  %   \end{columns}

    % sd %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[b]  
    \begin{column}{0.4\textwidth}
      \huge{\textbf{Stream Density:} the length of all streams (minus artificial lines through lakes) divided by the area of watershed.}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        % \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,ultra thick, cyan] (2,3) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (0.5,2.9) to [out=0,in=90] (1.1,1.9) to [out=270,in=90] (0,0);

        \draw [-,ultra thick, cyan] (-1,4) to [out=200,in=0] (-2,3) to [out=290,in=45] (-1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (-0.5,2.9) to [out=250,in=250] (-1.1,1.9) to [out=270,in=90] (0,0);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        %   \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,ultra thick, cyan] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

  % \rule{0.9\textwidth}{0.8pt}

  % sr %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[b]  
    \begin{column}{0.4\textwidth}
      \huge{\textbf{Stream Order Ratio:} The total number of first order stream reaches divided by the total number of higher order stream reaches. In this example the lower connectivity lake has a stream order ratio of 3 whereas the higher connectivity lake has a stream order ratio of 2.}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        % \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,ultra thick, cyan] (2,3) to [out=180,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,ultra thick, cyan] (0.5,2.9) to [out=0,in=90] (1.1,1.9);
        \draw [-, line width = 7pt, cyan] (1.1,1.9) to [out=270,in=90] (0,0);
        \draw [-,ultra thick, cyan] (1.8, 2.1) to [out=290,in=0] (1,1.5);

        \draw [-,ultra thick, cyan] (-1,4) to [out=200,in=0] (-2,3) to [out=290,in=135] (-1.1,1.9);
        \draw [-, line width = 7pt, cyan] (-1.1,1.9) to [out=270,in=90] (0,0);
        \draw [-,ultra thick, cyan] (-0.5,2.9) to [out=250,in=250] (-1.1,1.9) to [out=270,in=90] (0,0);
        \draw [-,ultra thick, cyan] (-1.8, 2.1) to [out=250,in=180] (-1,1.5);

        \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        %   \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,ultra thick, cyan] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-, ultra thick, cyan] (0.9,4) to [out=290,in=0] (1.1,1.9);
          \draw [-,line width=7pt, cyan] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw[cyan, ultra thick, domain=0:350, smooth cycle, fill=cyan!70!black!100] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

  \end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}