### 2018-07-12 figures/06_k.Rmd
order k figure by connectivity not metric value

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
# knitr::read_chunk("../scripts/03_viz.R")
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
# setwd("scripts")
# source("99_utils.R")
```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
  source("../scripts/01.5_loaddata.R")
```

```{r 06_k, echo=FALSE, message=FALSE, fig.cap="Distribution of the k parameter from the Vollenweider equation in low and high connectivity partitions at the (A) lake watershed and (B) network watershed scales. Green symbols indicate the lower of the two partition groups while purple symbols represent the higher of the two partition groups (see Table 2).", fig.width=7, fig.height=3.8, warning=FALSE}

fit_df_iws <- rbind(
  extract_coefs(readRDS("../data/global_vollenweider.rds"), "k"), 
  extract_coefs2(readRDS("../data/lc_vollenweider.rds"), 
                 c("k1_lc", "k2_lc")),
  extract_coefs2(readRDS("../data/md_vollenweider.rds"), 
                 c("k1_md", "k2_md")),
  extract_coefs2(readRDS("../data/iws/la_vollenweider.rds"),
                 c("k1_la", "k2_la")),
  extract_coefs2(readRDS("../data/iws/cd_vollenweider.rds"),
                 c("k1_cd", "k2_cd")),
  extract_coefs2(readRDS("../data/iws/ll_vollenweider.rds"),
                 c("k1_ll", "k2_ll")),
  extract_coefs2(readRDS("../data/iws/bf_vollenweider.rds"),
                 c("k1_bf", "k2_bf")),
  extract_coefs2(readRDS("../data/iws/sd_vollenweider.rds"),
                 c("k1_sd", "k2_sd")),
  extract_coefs2(readRDS("../data/iws/sr_vollenweider.rds"), 
                 c("k1_sr", "k2_sr")))
  #               extract_coefs2(fit_wc, c("k1_wc", "k2_wc")),

# controls ordering of ridge plot
fit_df_iws$key <- factor(fit_df_iws$key, levels = c("k",
                                                    "k1_sr", "k2_sr",
                                                    "k1_bf", "k2_bf",
                                                    "k2_md", "k1_md",
                                                    "k1_la", "k2_la",
                                                    "k2_lc", "k1_lc",
                                                    "k2_sd", "k1_sd",
                                                    "k1_cd", "k2_cd",
                                                    "k1_ll", "k2_ll"))
# k2  = higher metric value
# right = higher connectivity, purple

k_viz_iws <- ggplot(
  fit_df_iws, 
  aes(x = coef, y = key, fill = key, color = key)) + 
    geom_density_ridges(size = 0.1, scale = 0.7, rel_min_height = 0.01) + 
    geom_hline(yintercept = seq(1.8, 16.8, by = 2)) +
    scale_fill_manual(values = c("grey", 
                            rep(c(viridis::viridis(1, begin = 0.5), 
                                  viridis::viridis(1, begin = 0)), 8))) +
    scale_color_manual(values = c("grey", 
                               rep(c(viridis::viridis(1, begin = 0.5), 
                                     viridis::viridis(1, begin = 0)), 8))) +
    cowplot::theme_cowplot() + theme(panel.grid.major.y = element_blank(),
                                     axis.text = element_text(size = 10),
                                     axis.title = element_text(size = 12),
                               plot.title = element_text(size = 12, face = "bold", hjust = 0),
                                     axis.line.x = element_line(size = 1, 
                                     colour = "black"),
                                     axis.line.y = element_line(size = 1, 
                                                       colour = "black"), 
                                     legend.position = "none", 
                                     plot.margin = margin(0.5,0,0,0, "cm")) + 
  xlab("k Value") + ylab("") + 
  scale_y_discrete(breaks = levels(fit_df_iws$key)[
    c(1, seq(3, length(levels(fit_df_iws$key)), by = 2))], 
                   labels = rev(c("Link length",
                   "Closest lake distance",
                   "Stream density",
                   "Lake connection", 
                   "Upstream lake area",
                   "Max depth",
                   "Baseflow",
                   "Stream order ratio",
                   "Global")))

k_viz_iws

fit_df_nws <- rbind(
  extract_coefs(readRDS("../data/global_vollenweider.rds"), "k"), 
  extract_coefs2(readRDS("../data/lc_vollenweider.rds"), 
                 c("k1_lc", "k2_lc")),
  extract_coefs2(readRDS("../data/nws/cd_vollenweider.rds"), 
                 c("k1_cd", "k2_cd")),
  extract_coefs2(readRDS("../data/nws/ll_vollenweider.rds"), 
                 c("k1_ll", "k2_ll")),
  extract_coefs2(readRDS("../data/nws/la_vollenweider.rds"), 
                 c("k1_la", "k2_la")),
  extract_coefs2(readRDS("../data/nws/la_vollenweider.rds"), 
                 c("k1_la", "k2_la")),
  extract_coefs2(readRDS("../data/md_vollenweider.rds"), 
                 c("k1_md", "k2_md")),
  extract_coefs2(readRDS("../data/nws/bf_vollenweider.rds"), 
                 c("k1_bf", "k2_bf")),
  extract_coefs2(readRDS("../data/nws/sd_vollenweider.rds"), 
                 c("k1_sd", "k2_sd")),
  extract_coefs2(readRDS("../data/nws/sr_vollenweider.rds"), 
                 c("k1_sr", "k2_sr")))

# align iws and nws panel rows
fit_df_nws <- dplyr::filter(fit_df_nws, fit_df_nws$key %in% fit_df_iws$key)

fit_df_nws$key <- factor(fit_df_nws$key, levels = c("k",
                                                    "k1_sr", "k2_sr",
                                                    "k1_bf", "k2_bf",
                                                    "k2_md", "k1_md",
                                                    "k1_la", "k2_la",
                                                    "k2_lc", "k1_lc",
                                                    "k2_sd", "k1_sd",
                                                    "k1_cd", "k2_cd",
                                                    "k1_ll", "k2_ll"))
# k2  = higher metric value
# right = higher connectivity, purple

k_viz_nws <- ggplot(
  fit_df_nws, 
  aes(x = coef, y = key, fill = key, color = key)) + 
  geom_density_ridges(size = 0.1, scale = 0.7, rel_min_height = 0.01) + 
  geom_hline(yintercept = seq(1.8, 16.8, by = 2)) +
  scale_fill_manual(values = c("grey",
                               rep(c(viridis::viridis(1, begin = 0.5),
                                     viridis::viridis(1, begin = 0)), 8))) +
  scale_color_manual(values = c("grey", 
                               rep(c(viridis::viridis(1, begin = 0.5), 
                                     viridis::viridis(1, begin = 0)), 8))) +
  cowplot::theme_cowplot() + theme(axis.text = element_text(size = 10),
                                   axis.title = element_text(size = 12),
                               plot.title = element_text(size = 12, face = "bold", hjust = 0),
                                   axis.line.x = element_line(size = 1, 
                                                  colour = "black"),
                                   axis.line.y = element_line(size = 1, 
                                                  colour = "black"), 
                                   legend.position = "none", 
                                   plot.margin = margin(0.5,0,0,-0.7, "cm")) + 
  xlab("k Value") +  ylab("") +
  scale_y_discrete(breaks = levels(fit_df_nws$key)[
    c(1, seq(3, length(levels(fit_df_nws$key)), by = 2))], 
    labels = rev(c("Link length",
                   "Closest lake distance",
                   "Stream density",
                   "Lake connection", 
                   "Upstream lake area",
                   "Max depth",
                   "Baseflow",
                   "Stream order ratio",
                   "Global")))

k_viz_nws

plot_grid(
  k_viz_iws + ggtitle("A. Lake Watershed") + xlim(0.8, 1.6),
  k_viz_nws + ggtitle("B. Network Watershed") + xlim(0.8, 1.6), 
  ncol = 2, align = "v", rel_widths = c(1, 1)) + 
  theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))

```