### 2018-09-18 manuscript/figures.Rmd
align fig 1 dashes and symbols with caption

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

```{r fig.cap="Figure 1: Major lake and watershed P factors affecting lake P retention. Shaded symbols indicate tangible factors that are typically considered in P retention models while open symbols indicate additional factors specific to the present study. Dashed lines indicate inferred linkages, which cannot be tested with available data, but are otherwise discussed herein. The abbreviation ‘W. Res. Time’ is water residence time.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../figures/01_conceptual_p-cycle_pt2.pdf")
```

```{r fig.cap="Connectivity metric definitions along with examples of high and low connectivity lakes.", echo=FALSE}
knitr::include_graphics("../figures/02_conny_metric_key_crop_small.pdf")
```

```{r fig.cap="Diagram showing the lake watershed (LWS) and network watershed (NWS) of three lakes. Here the IWS of lake 3 encompasses the LWS of lake 2 because of it is smaller than 10 ha small size but it does not encompass the LWS of lake 1 because it has an area of at least 10 ha.  In contrast to the LWS boundaries, the NWS boundaries extend to the headwaters of the lake chain.", echo=FALSE}
knitr::include_graphics("../figures/03_iws_nws.pdf")
```

```{r 07_cor_mat_hmap, echo=FALSE, warning=FALSE, fig.cap="Correlation between connectivity metrics and selected lake characteristics.", message=FALSE}
knitr::include_graphics("../figures/07_cor_mat_hmap-1.pdf")
```

```{r 08_maps, echo=FALSE, fig.cap="Maps showing the locations of lake connectivity partitions. Green symbols indicate the lower of the two partition groups while purple symbols represent the higher of the two partition groups (see Table 2).", warning=FALSE, message=FALSE}
knitr::include_graphics("../figures/08_maps-1.pdf")
```

```{r 04_global_vollenweider, echo=FALSE, fig.cap="Residence time (yr) versus P retention for the NES dataset and the global model fit to the data where the solid line and shaded interval represents the median and central 95% interval estimates respectively.", warning=FALSE}

knitr::include_graphics("../figures/04_global_vollenweider-1.pdf")
```

```{r 06_k, echo=FALSE, message=FALSE, fig.cap="Distribution of the k parameter from the Vollenweider's equation in low and high connectivity partitions at the (A) LWS and (B) NWS scales. Green symbols indicate the lower of the two partition groups while purple symbols represent the higher of the two partition groups (see Table 2).", fig.width=5, fig.height=3.8, warning=FALSE}
knitr::include_graphics("../figures/06_k-3.pdf")
```