### 2018-10-15 figures/06_k.Rmd
change network to lake network

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
# knitr::read_chunk("../scripts/03_viz.R")
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
# setwd("scripts")
# source("99_utils.R")
```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
  source("../scripts/01.5_loaddata.R")
```

```{r 06_k, echo=FALSE, message=FALSE, fig.cap="Distribution of the k parameter from the Vollenweider equation in low and high connectivity partitions at the (A) lake watershed and (B) network watershed scales. Green symbols indicate the lower of the two partition groups while purple symbols represent the higher of the two partition groups (see Table 2).", fig.width=7, fig.height=3.8, warning=FALSE}

model_signif <- read.csv("../scripts/model_signif.csv", 
                         stringsAsFactors = FALSE)

fit_df_iws <- rbind(
  extract_coefs(readRDS("../data/global_vollenweider.rds"), "k"), 
  extract_coefs2(readRDS("../data/lc_vollenweider.rds"), 
                 c("k1_lc", "k2_lc")),
  extract_coefs2(readRDS("../data/md_vollenweider.rds"), 
                 c("k1_md", "k2_md")),
  extract_coefs2(readRDS("../data/iws/la_vollenweider.rds"),
                 c("k1_la", "k2_la")),
  extract_coefs2(readRDS("../data/iws/cd_vollenweider.rds"),
                 c("k1_cd", "k2_cd")),
  extract_coefs2(readRDS("../data/iws/ll_vollenweider.rds"),
                 c("k1_ll", "k2_ll")),
  extract_coefs2(readRDS("../data/iws/bf_vollenweider.rds"),
                 c("k1_bf", "k2_bf")),
  extract_coefs2(readRDS("../data/iws/sd_vollenweider.rds"),
                 c("k1_sd", "k2_sd")),
  extract_coefs2(readRDS("../data/iws/sr_vollenweider.rds"), 
                 c("k1_sr", "k2_sr")))
  #               extract_coefs2(fit_wc, c("k1_wc", "k2_wc")),

# controls ordering of ridge plot
fit_df_iws$key <- factor(fit_df_iws$key, levels = c("k",
                                                    "k1_sr", "k2_sr",
                                                    "k1_bf", "k2_bf",
                                                    "k2_md", "k1_md",
                                                    "k1_la", "k2_la",
                                                    "k2_lc", "k1_lc",
                                                    "k2_sd", "k1_sd",
                                                    "k1_cd", "k2_cd",
                                                    "k1_ll", "k2_ll"))
# k2  = higher metric value
# right = higher connectivity, purple

iws_labels_raw <- data.frame(parameter = rev(c(
                   "Link length",
                   "Closest lake distance",
                   "Stream density",
                   "Lake connection", 
                   "Upstream lake area",
                   "Max depth",
                   "Baseflow",
                   "Stream order ratio",
                   "Global")), stringsAsFactors = FALSE) %>%
  mutate(parameter_alt = parameter)
iws_labels_raw$parameter_alt[iws_labels_raw$parameter == "Link length"] <- 
  "Average link length"

colorado <- function(src) {
  # https://stackoverflow.com/a/39695859/3362993
  src_levels <- src                                 # retrieve the levels in their order
  boulder <- src_levels[grep("\\*", src_levels)]
  brave <- seq_len(length(src_levels)) %in% grep("\\*", src_levels)                       # make sure everything we want to make bold is actually in the factor levels
  b_pos <- purrr::map_int(boulder, ~which(.==src_levels)) # then find out where they are
  b_vec <- rep("plain", length(src_levels))               # make'm all plain first
  b_vec[b_pos] <- "bold"                                  # make our targets bold
  b_vec                                                   # return the new vector
}

iws_labels <- left_join(iws_labels_raw, model_signif, 
                        by = c("parameter_alt" = "parameter")) %>%
  filter(scale == "iws" | is.na(scale)) %>%
  arrange(match(parameter, iws_labels_raw$parameter)) %>%
  mutate(signif = sapply(signif, signif_star)) %>%
  mutate(parameter = paste0(parameter, signif)) %>%
  select(parameter)

nws_labels <- left_join(iws_labels_raw, model_signif, 
                        by = c("parameter_alt" = "parameter")) %>%
  filter(scale == "nws" | is.na(scale)) %>%
  arrange(match(parameter, iws_labels_raw$parameter)) %>%
  mutate(signif = sapply(signif, signif_star)) %>%
  mutate(parameter = paste0(parameter, signif)) %>%
  select(parameter)

k_viz_iws <- ggplot(
  fit_df_iws, 
  aes(x = coef, y = key, fill = key, color = key)) + 
    geom_density_ridges(size = 0.1, scale = 0.7, rel_min_height = 0.01) + 
    geom_hline(yintercept = seq(1.8, 16.8, by = 2)) +
    scale_fill_manual(values = c("grey", 
                            rep(c(viridis::viridis(1, begin = 0.5), 
                                  viridis::viridis(1, begin = 0)), 8))) +
    scale_color_manual(values = c("grey", 
                               rep(c(viridis::viridis(1, begin = 0.5), 
                                     viridis::viridis(1, begin = 0)), 8))) +
    cowplot::theme_cowplot() + theme(panel.grid.major.y = element_blank(),
                                     axis.text = element_text(size = 9),
                                     axis.text.y = element_text(size = 8, 
                                                face = colorado(iws_labels$parameter)),
                                     axis.title = element_text(size = 10),
                               plot.title = element_text(size = 9, face = "bold", hjust = 0),
                                     axis.line.x = element_line(size = 1, 
                                     colour = "black"),
                                     axis.line.y = element_line(size = 1, 
                                                       colour = "black"), 
                                     legend.position = "none", 
                                     plot.margin = margin(0.5,0,0,0, "cm")) + 
  xlab("k Value") + ylab("") + 
  scale_y_discrete(breaks = levels(fit_df_iws$key)[
    c(1, seq(3, length(levels(fit_df_iws$key)), by = 2))], 
                   labels = iws_labels$parameter)

k_viz_iws

fit_df_nws <- rbind(
  extract_coefs(readRDS("../data/global_vollenweider.rds"), "k"), 
  extract_coefs2(readRDS("../data/lc_vollenweider.rds"), 
                 c("k1_lc", "k2_lc")),
  extract_coefs2(readRDS("../data/nws/cd_vollenweider.rds"), 
                 c("k1_cd", "k2_cd")),
  extract_coefs2(readRDS("../data/nws/ll_vollenweider.rds"), 
                 c("k1_ll", "k2_ll")),
  extract_coefs2(readRDS("../data/nws/la_vollenweider.rds"), 
                 c("k1_la", "k2_la")),
  extract_coefs2(readRDS("../data/nws/la_vollenweider.rds"), 
                 c("k1_la", "k2_la")),
  extract_coefs2(readRDS("../data/md_vollenweider.rds"), 
                 c("k1_md", "k2_md")),
  extract_coefs2(readRDS("../data/nws/bf_vollenweider.rds"), 
                 c("k1_bf", "k2_bf")),
  extract_coefs2(readRDS("../data/nws/sd_vollenweider.rds"), 
                 c("k1_sd", "k2_sd")),
  extract_coefs2(readRDS("../data/nws/sr_vollenweider.rds"), 
                 c("k1_sr", "k2_sr")))

# align iws and nws panel rows
fit_df_nws <- dplyr::filter(fit_df_nws, fit_df_nws$key %in% fit_df_iws$key)

fit_df_nws$key <- factor(fit_df_nws$key, levels = c("k",
                                                    "k1_sr", "k2_sr",
                                                    "k1_bf", "k2_bf",
                                                    "k2_md", "k1_md",
                                                    "k1_la", "k2_la",
                                                    "k2_lc", "k1_lc",
                                                    "k2_sd", "k1_sd",
                                                    "k1_cd", "k2_cd",
                                                    "k1_ll", "k2_ll"))
# k2  = higher metric value
# right = higher connectivity, purple

k_viz_nws <- ggplot(
  fit_df_nws, 
  aes(x = coef, y = key, fill = key, color = key)) + 
  geom_density_ridges(size = 0.1, scale = 0.7, rel_min_height = 0.01) + 
  geom_hline(yintercept = seq(1.8, 16.8, by = 2)) +
  scale_fill_manual(values = c("grey",
                               rep(c(viridis::viridis(1, begin = 0.5),
                                     viridis::viridis(1, begin = 0)), 8))) +
  scale_color_manual(values = c("grey", 
                               rep(c(viridis::viridis(1, begin = 0.5), 
                                     viridis::viridis(1, begin = 0)), 8))) +
  cowplot::theme_cowplot() + theme(axis.text = element_text(size = 9),
                                   axis.text.y = element_text(size = 8, 
                                                face = colorado(nws_labels$parameter)),
                                   axis.title = element_text(size = 10),
                               plot.title = element_text(size = 9, face = "bold", hjust = 0),
                                   axis.line.x = element_line(size = 1, 
                                                  colour = "black"),
                                   axis.line.y = element_line(size = 1, 
                                                  colour = "black"), 
                                   legend.position = "none", 
                                   plot.margin = margin(0.5,0,0,-0.7, "cm")) + 
  xlab("k Value") +  ylab("") +
  scale_y_discrete(breaks = levels(fit_df_nws$key)[
    c(1, seq(3, length(levels(fit_df_nws$key)), by = 2))], 
    labels = nws_labels$parameter)

k_viz_nws

plot_grid(
  k_viz_iws + ggtitle("a. Lake Watershed") + xlim(0.8, 1.6),
  k_viz_nws + ggtitle("b. Lake Network Watershed") + xlim(0.8, 1.6), 
  ncol = 2, align = "v", rel_widths = c(1, 1)) + 
  theme(plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"))

```