### 2018-10-17 manuscript/appendix.Rmd
cleanup appendix, placeholder for milstead text

---
title: "Appendix"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 5)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
# setwd("scripts")
# source("99_utils.R")
```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
  if(file.exists("../scripts/01.5_loaddata.R")){
    source("../scripts/01.5_loaddata.R")  
  }else{
    source("scripts/01.5_loaddata.R")  
  }
```

```{r read_metric_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../data/connectivity_metrics.R")
```

```{r load_packages, echo=FALSE, warning=FALSE, message=FALSE}

```

```{r source_functions, echo=FALSE, warning=FALSE, message=FALSE}

```

milstead reanalysis

```{r milstead_reanalysis, fig.cap="Relationship between residence time and phosphorus retention for inland lakes with surface area greater than 4 ha in Northeastern USA (New England). Vertical dashed lines denote lakes with intermediate residence times (within the interquartile range). Best fit lines are colored based on whether they have upstream lake connections (DR_LakeStream), or upstream stream connection (DR_Stream), are headwater lakes, or are isolated lakes. Data from Milstead et al. (2013).", fig.width = 5, fig.height = 4, warning=FALSE, message=FALSE}
library(cowplot)
library(ggplot2)
library(sf)
library(nhdR)
library(LAGOSNE)
library(maps)
library(dplyr)

## load data ####
milstead_path <- "../data/milstead_2013.csv"
# unlink(milstead_path)

if(!file.exists(milstead_path)){
  states <- sf::st_as_sf(map("state", plot = FALSE, fill = TRUE))
  nhd_plus <- nhdR::nhd_plus_load(c(1,2), dsn = "NHDWaterbody")
  nhd_plus <- st_transform(nhd_plus, st_crs(states))
  states <- states[which(sapply(
             sf::st_covers(states, nhd_plus),
             function(x) length(x)) > 0),]

  nhd <- nhdR::nhd_load(state.abb[tolower(state.name) %in% states$ID], "NHDWaterbody")
  nhd$GNIS_ID <- sapply(as.character(nhd$GNIS_ID), function(x) substring(x, 3, nchar(x)))
  nhd <- nhd[!is.na(nhd$ReachCode),]
  st_geometry(nhd) <- NULL
  lg <- lagosne_load("1.087.1")

  library(fulltext)
  query <- ft_search(query = "Estimating Summer Nutrient Concentrations in
                      Northeastern Lakes from SPARROW Load Predictions and Modeled
                      Lake Depth and Volume", from = "plos")
  doi <- ft_links(query)$plos$ids
  milstead <- read.csv(ft_get_si(doi, 1))

  # paused here
  milstead_j <- dplyr::left_join(milstead, 
                  dplyr::select(nhd_plus, COMID, REACHCODE), 
                by = c("WB_ID" = "COMID")) %>%
    dplyr::left_join(
        dplyr::select(nhd, Permanent_Identifier, ReachCode), 
      by = c("REACHCODE" = "ReachCode")) %>%
    dplyr::left_join(
        dplyr::select(lg$lakes.geo, lakeconnection, lakes_nhdid), 
      by = c("Permanent_Identifier" = "lakes_nhdid")) %>%
    dplyr::left_join(
        dplyr::select(lg$locus, nhdid, nhd_lat, nhd_long), 
      by = c("Permanent_Identifier" = "nhdid"))
  milstead_j <- milstead_j[complete.cases(milstead_j[,c("nhd_lat", "nhd_long")]),]
  milstead_j <- milstead_j[,nchar(names(milstead_j)) < 30] 
  milstead_j <- milstead_j[, -which(names(milstead_j) == "geometry")]

  write.csv(milstead_j, milstead_path, row.names = FALSE)
}

md    <- read.csv(milstead_path, stringsAsFactors = FALSE)
md$Rp <- 1 - (as.numeric(md$Poutput) / as.numeric(md$Pinput))
md    <- md[md$Rp >= 0 & md$Rp <= 1,]
md$hrt <- as.numeric(md$hrt)
md     <- md[md$hrt > 0,] 
md     <- md[md$Area > 0.04,] # rm less than 4 ha

# library(magrittr)
# glm.fit <- milstead %>% dplyr::group_by(lakeconnectivity) %>% dplyr::do(fit = glm(Rp ~ hrt, family = "binomial", data = .))
# 
# broom::tidy(glm.fit, fit)
# glm.stats <- as.data.frame(broom::glance(glm.fit, fit))
# # https://ww2.coastal.edu/kingw/statistics/R-tutorials/logistic.html
# 1 - pchisq(glm.stats$null.deviance - glm.stats$deviance, 1)

# ggplot(data = milstead[!is.na(milstead$lakeconnectivity),], aes(x = lakeconnectivity, y = hrt)) + geom_boxplot(notch = TRUE) + scale_y_log10()

yr_labels <- c("1e-07", "1e-05", 
               as.character(10/365), 
               as.character(2/12), as.character(6/12), 
               "2", "10")

format_labels <- function(x, mult_factor){
  gsub("\\.+$", "\\1", gsub("0+$", "\\1",
    trimws(format(round(as.numeric(x), 6) * 
      mult_factor, scientific = FALSE)), perl = TRUE))
  }

day_labels   <- ceiling(as.numeric(
  format_labels(yr_labels, 365)) / 10) * 10
minute_labels  <- ceiling(as.numeric(
  format_labels(yr_labels, 365 * 24 * 60)) / 10) * 10
month_labels  <- round(as.numeric(
  format_labels(yr_labels, 12)), 0)

mixed_labels <- paste0(
      c(minute_labels[1:2], day_labels[3], month_labels[4:5], yr_labels[6:7]), 
      c(" min", " min", " days", " months", " months", " years", " years"))

quants <- exp(quantile(log(md[, "hrt"]), na.rm = TRUE))

gg_fit_single <- 
  ggplot(data = md, 
    aes(x = hrt, y = Rp)) + geom_point(size = 0.9) +
  scale_x_log10(labels = mixed_labels, 
    breaks = as.numeric(yr_labels), limits = c(1 / 365, 11)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              se = TRUE) +
  scale_color_brewer(palette = "Set1") +
  cowplot::theme_cowplot() + 
  theme(legend.position = "none", legend.title = element_blank(), legend.text = element_text()) +
  xlab("Residence Time") +
  ylab("P Retention (%)") + 
  geom_segment(data = data.frame(x = quants[c(2, 4)], y = c(0.8, 0.8)), 
               aes(x = x, y = c(0, 0), xend = x, yend = y), 
               color = "gray42", size = 1.5, linetype = 2)

# if(!interactive()){
#   ggsave("milstead_single.pdf", gg_fit_single, height = 5)
# }

# milstead <- droplevels(milstead[milstead$lakeconnectivity != "Isolated",])
gg_fit_multi <- ggplot(data = md, aes(x = hrt, y = Rp)) +
  geom_point(size = 0.9) +
  scale_x_log10(labels = mixed_labels, 
                breaks = as.numeric(yr_labels), limits = c(1 / 365, 11)) +
  stat_smooth(method = "glm", method.args = list(family = "binomial"), 
              se = TRUE, aes(color = lakeconnection), 
              data = md) +
  scale_color_brewer(palette = "Set1") +
  cowplot::theme_cowplot() + 
  theme(legend.title = element_blank(), legend.position = c(0.13, 0.8),
        legend.text = element_text(), legend.background = element_rect(fill = "white")) +
  xlab("Residence Time") +
  ylab("P Retention (%)") + 
  geom_segment(data = data.frame(x = quants[c(2, 4)], y = c(0.8, 0.8)), 
               aes(x = x, y = c(0, 0), xend = x, yend = y), 
               color = "gray42", size = 1.5, linetype = 2)

gg_fit_multi
```

```{r p-retention_vs_percent-ag, fig.cap="P retention versus percent agriculture in network watersheds.", fig.width = 3, fig.height = 3, warning=FALSE, message=FALSE}
nes_nws$percent_ag <- nes_iws$iws_nlcd1992_pct_81 + nes_iws$iws_nlcd1992_pct_82

ggplot() + 
  geom_point(data = nes_nws, aes(x = percent_ag, y = p_percent_retention)) +
  xlab("Percent Agriculture") + ylab("P Retention")

```

```{r nes_vs_nla, fig.cap="Comparison of selected lake characteristics among the stratified random sampling design of the National Lakes Assessment (nla) and the haphazard sampling of the National Eutrophication Survey (nes) lakes analyzed in the present study.", fig.height=4, warning=FALSE, message=FALSE}
# , message=FALSE, warning=FALSE
nla <- nla_load(2012)

tp <- select(nla$waterchem_wide, PTL_RESULT, UID) %>%
      left_join(select(nla$wide_siteinfo, AREA_HA, SITE_ID, UID)) %>%
      group_by(SITE_ID) %>%
      summarize(tp = median(PTL_RESULT), area = median(AREA_HA) / 100)

chl  <- select(nla$chla_wide, CHLX_RESULT, UID)  %>%
        left_join(select(nla$wide_siteinfo, SITE_ID, UID)) %>%
        group_by(SITE_ID) %>%
        summarize(chl = median(CHLX_RESULT))

secchi  <- select(nla$secchi, SECCHI, SITE_ID)  %>%
           group_by(SITE_ID) %>%
           summarize(secchi = median(SECCHI))

depth <- select(nla$wide_phab, DEPTH_AT_STATION, SITE_ID)  %>%
           group_by(SITE_ID) %>%
           summarize(depth = median(DEPTH_AT_STATION))

nla <- reduce(list(tp, chl, secchi, depth), left_join)
# skimr::skim(res)
nla <- gather(select(nla, -SITE_ID), variable, value)
nla$id <- "nla"

nes <- select(nes_iws, tp, area = surface_area, chl, secchi, depth = maxdepth)
nes$tp <- 1000 * nes$tp # mg to ug
nes <- gather(nes, variable, value)
nes$id <- "nes"

res <- rbind(nla, nes)

# ggplot(data = dplyr::filter(res, variable == "chl"), aes(y = id, x = value + 0.1)) + 
#   geom_density_ridges() + 
#   scale_x_log10()

ggplot(data = res, aes(y = id, x = value + 0.1)) + 
  geom_density_ridges() + 
  scale_x_log10() + 
  facet_wrap(~variable) + 
  theme(axis.title = element_blank())
```