### 2018-12-19 figures/02_conny_metric_key.tex
recolor metric key

\documentclass[bigger]{beamer}
\usepackage[orientation=portrait, size=a10, scale=1]{beamerposter}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{tabularx}

\setbeamertemplate{navigation symbols}{}%remove navigation symbols
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\input{beamer-lake-fig/beamer-lake-fig.tex}

\definecolor{vhigh}{HTML}{440154}
\definecolor{vlow}{HTML}{21908C}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Connectivity Metric Key}
\begin{frame}
  \begin{center}

% labels
    \begin{columns}[b]     
      \begin{column}{0.4\textwidth}        
      \end{column}

      \begin{column}{0.25\textwidth}
        \huge{Low}
      \end{column}

      \begin{column}{0.25\textwidth}
        \huge{High}
      \end{column}
   \end{columns}

% spacer  
  \begin{columns}    
    \begin{column}[b]{0.4\textwidth}
      \vspace{2em}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}
  \end{columns}

% lake label  
  \begin{columns}[T]    
    \begin{column}{0.4\textwidth}
      {\fontsize{65}{65}\textbf{Lake-based Metrics}}
      \vspace{3em}
    \end{column}

  \begin{column}{0.25\textwidth}
  \end{column}

  \begin{column}{0.25\textwidth}
  \end{column}
  \end{columns}

  % cd %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[T]    
    \begin{column}{0.4\textwidth}
      \huge{\textbf{A. Closest Lake Distance:} Network distance to the closest upstream lake.}
    \end{column}

  \begin{column}{0.25\textwidth}
    \begin{tikzpicture}[scale = 2]
    %\draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,line width = 4pt, vlow] (2.0,2.5) to [out=270,in=90] (0.2,0);
        \draw[vlow, ultra thick, domain=0:350, smooth cycle, fill=vlow, yshift = 3cm, xshift = 2cm] plot (\x:0.6+rnd*0.2);
        \draw[vlow!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vlow] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
  \end{column}

  \begin{column}{0.25\textwidth}
  \vspace{2em}
    \begin{tikzpicture}[scale = 2]
        %\draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw[-,line width = 4pt, vhigh] plot [smooth, tension = 1] coordinates {(2.2,2.8) (1.3,1.6) (-0.3,2.7) (0,0)};

        \draw[vhigh, ultra thick, domain=0:350, smooth cycle, fill=vhigh, yshift = 3cm, xshift = 2cm] plot (\x:0.6+rnd*0.2);
        \draw[vhigh!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vhigh] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
  \end{column}

  \end{columns}

% spacer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{columns}    
    \begin{column}[b]{0.4\textwidth}
    \vspace{2em}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}
  \end{columns}

  % la %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[T]
    \begin{column}{0.4\textwidth}
      \huge{\textbf{B. Upstream Lake Area:} Total area of upstream lakes limited to lakes that are greater than or equal to 4ha.}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
          \draw [-,line width = 4pt, vlow] (2.2,3.7) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw[vlow, ultra thick, domain=0:350, smooth cycle, fill=vlow, yshift = 3.5cm, xshift = 2cm] plot (\x:0.5+rnd*0.2);
          \draw[vlow!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vlow] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
          \draw [-,line width = 4pt, vhigh] (2.2,4.1) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw[vhigh, ultra thick, domain=0:350, smooth cycle, fill=vhigh, yshift = 3.5cm, xshift = 2cm] plot (\x:0.9+rnd*0.2);
          \draw[vhigh!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vhigh] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

\begin{columns}    
    \begin{column}[b]{0.4\textwidth}
    \vspace{4em}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}
  \end{columns}

\begin{columns}    
    \begin{column}[b]{0.4\textwidth}
      {\fontsize{65}{65}\textbf{Lake-stream-based Metrics}}
      \vspace{2em}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}
  \end{columns}

  % ll %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[T] 
    \begin{column}{0.4\textwidth}
      \huge{\textbf{C . Average Link Length:} Sum of the total length of stream reaches between junctions divided by the total number of reaches.}
    \end{column}

    \begin{column}{0.25\textwidth}
      \vspace{3em}
      \begin{tikzpicture}[scale = 2]
        %\draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,line width = 4pt, vlow] (1.6,2.4) to [out=180,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,line width = 4pt, vlow] (0.58,2.3) to [out=0,in=90] (1.1,1.9) to [out=270,in=90] (0,0);

        \draw [-,line width = 4pt, vlow] (-1.2,2.75) to [out=200,in=0] (-1.6,2.4) to [out=290,in=110] (-1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,line width = 4pt, vlow] (-0.58,2.3) to [out=250,in=70] (-1.1,1.9) to [out=270,in=90] (0,0);

        \draw[vlow!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vlow] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \begin{column}{0.25\textwidth}
      \vspace{-3em}
      \begin{tikzpicture}[scale = 2]
           %\draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,line width = 4pt, vhigh] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,line width = 4pt, vhigh] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw [-,line width = 4pt, vhigh] (-2.2,4.8) to [out=90,in=180] (-1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,line width = 4pt, vhigh] (-0.9,4) to [out=260,in=180] (-1.1,1.9) to [out=270,in=90] (0,0);

          \draw[vhigh!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vhigh] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}
    \end{columns}

    % spacer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \begin{columns}    
    \begin{column}[b]{0.4\textwidth}
    \vspace{2em}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}
  \end{columns}

    % sd %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[T]  
    \begin{column}{0.4\textwidth}
      \huge{\textbf{D. Stream Density:} the length of all streams (minus artificial lines through lakes) divided by the area of watershed.}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        %   \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw[vlow!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vlow] plot (\x:1+rnd*0.5);

        \draw [-,line width = 4pt, vlow] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
        \draw [-,line width = 4pt, vlow] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
        \end{tikzpicture}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
          %\draw[step=1cm,gray,very thin] (-4,-4) grid (4,4);
          \draw [-,line width = 4pt, vhigh] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-,line width = 4pt, vhigh] (0.9,4) to [out=290,in=0] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw [-,line width = 4pt, vhigh] (2.9,3.4) to [out=290,in=40] (1.95,3);
          \draw [-,line width = 4pt, vhigh] (0.2,3.9) to [out=310,in=190] (1.25,3.2);
          \draw [-,line width = 4pt, vhigh] (1.6,4.8) to [out=290,in=85] (2.07,3.6);

          \draw[vhigh!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vhigh] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}

    \end{columns}

  % \rule{0.9\textwidth}{0.8pt}

  % spacer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \begin{columns}    
    \begin{column}[b]{0.4\textwidth}
    \vspace{2em}
    \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}

  \begin{column}[b]{0.25\textwidth}
  \end{column}
  \end{columns}

  % sr %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \begin{columns}[T]  
    \begin{column}{0.4\textwidth}
      \huge{\textbf{E. Stream Order Ratio:} The total number of first order stream reaches divided by the total number of higher order stream reaches. In this example the lower value lake has a stream order ratio of 2 whereas the higher value lake has a stream order ratio of 3.}
    \end{column}

    \begin{column}{0.25\textwidth}
      \begin{tikzpicture}[scale = 2]
        %   \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
          \draw [-,line width = 4pt, vlow] (2.2,4.8) to [out=90,in=0] (1.1,1.9) to [out=270,in=90] (0,0);
          \draw [-, line width = 4pt, vlow] (0.9,4) to [out=290,in=0] (1.1,1.9);
          \draw [-,line width=4pt, vlow] (1.1,1.9) to [out=270,in=90] (0,0);

          \draw[vlow!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vlow] plot (\x:1+rnd*0.5);
        \end{tikzpicture}
    \end{column}

    \begin{column}{0.25\textwidth}
    %\vspace{-2em}
    \vspace{4em}
      \begin{tikzpicture}[scale = 2]
        % \draw[step=1cm,gray,very thin] (-4,-3) grid (4,4);
        \draw [-,line width = 4pt, vhigh] (2,3) to [out=180,in=0] (1.1,1.9) to [out=270,in=90] (0,0);        
        \draw [-,line width = 4pt, vhigh] (0.5,2.9) to [out=0,in=90] (1.1,1.9);
        \draw [-, line width = 8pt, vhigh] (1.1,1.9) to [out=270,in=90] (0,0);
        \draw [-,line width = 4pt, vhigh] (1.8, 2.1) to [out=290,in=0] (1,1.5);

        \draw [-,line width = 4pt, vhigh] (-1,4) to [out=200,in=0] (-2,3) to [out=290,in=135] (-1.1,1.9);
        \draw [-, line width = 8pt, vhigh] (-1.1,1.9) to [out=250,in=90] (0,0);
        \draw [-,line width = 4pt, vhigh] (-0.5,2.9) to [out=250,in=250] (-1.06,1.95) to [out=250,in=70] (0,0);
        \draw [-,line width = 4pt, vhigh] (-1.8, 2.1) to [out=250,in=180] (-1,1.5);

        \draw[vhigh!70!black!100, line width = 4pt, domain=0:350, smooth cycle, fill=vhigh] plot (\x:1+rnd*0.5);
      \end{tikzpicture}
    \end{column}

    \end{columns}

  \end{center}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}